---
title: "Exercise 2"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




## install igraph

```{r }
install.packages("igraph", repos = "http://cran.us.r-project.org")
library(igraph)
```


## Load libraries

```{r }
library(tidyverse)
library(tidygraph)
library(ggraph)
library(readr)

```

## Load nodes table

```{r }
nodes <- read_csv("Ex2 - NODES.csv")
head(nodes)
```

## Load edges table

```{r }
edges <- read_csv("Ex2 - EDGES.csv")
head(edges)
```


## Plot network

Create object that ggraph could work with

```{r }
ig <- igraph::graph_from_data_frame(edges, vertices = nodes) %>% as_tbl_graph()

```

Create and display plot with lables

```{r }
social_net <- ggraph(ig, layout = "stress") +                                                                                                         
  geom_node_point(size = 2) +                                         
  geom_node_text(aes(label = name), nudge_y = 0.05, nudge_x = 0.2)+ 
  geom_edge_link() +
  theme_void()

show(social_net)
```

From this plot, degree centrality puts B, C, D and 3 as most central nodes.
However A seems to be node with highest betweenness centrality

## Determine measures of centrality


1- fonction
2- resultat (table)
3- plot with degree centrality
4 - interpretation

for each of the 3 measures of centrality

Than final discussion



- Degree centrality:

"The simplest measure of centrality is degree centrality. It counts how many edges each node has - the most degree central actor is the one with the most ties." (https://bookdown.org/markhoff/social_network_analysis/centrality.html)

```{r }
degree <- degree(ig)
degree
```

This table confirms the same level of degree centrality for B, C, D and Three. This is not very helfpul if we are to pick one seat in the bus, B, C and D would have the same interest for creating informal social connexions. Only A can be eliminated. We would need further attributes to distinguish the advantages of B, C, and D so as to pick one seat to sit, or look at different centrality measure.


- Closeness:
We measure the distance between two nodes as the length of the shortest path between them. Farness, for a given node, is the average distance from that node to all other nodes. Closeness is then the reciprocal of farness (1/farness).

```{r }
closeness <- closeness(ig)
closeness
```
I do not understand why R would not acheive to find a value for the closeness measure for Four and Six


- Betweenness centrality

betweenness effectively counts how many shortest paths each node is on. The higher a node’s betweenness, the more important they are for the efficient flow of goods in a network.

```{r }
betweenness <- betweenness(ig)
betweenness
```
From betweenness centrality, we find that A is the best choice to pick a sit because it is the node that concentrates the flow of information in the network. This appears in the plot of the network we did earlier.


Overall, degree centrality and betweenness give different indications on which seat to pick.
We would need to calculate other measures for centrality, maybe with directionnal edges, or at least have more information about each node (attributes of the node).


## Plot with centrality degree


```{r }

library(graphlayouts)

ggraph(ig, "stress", bbox = 15) +
  geom_edge_link2(aes(edge_colour = ""), edge_width = 0.5) +
  geom_node_point(aes(fill = betweenness, size = degree), shape = 21) +
  geom_node_text(aes(label = name, size=closeness),
    family = "mono", repel = TRUE
  ) +
  scale_edge_colour_brewer(palette = "Set1") +
  scale_size(range = c(2, 5), guide = "none") +
  theme_graph() 


```
